\onecolumn
	%% copyrightpage
	\begingroup
	\footnotesize
	\parindent 0pt
	\parskip \baselineskip
	\textcopyright{} 2022 Antonius Torode \\
	All rights reserved.
	
	The articles of this work may be distributed freely. 
	
	
	\textbf{About this document:} The Peer Reviewed pirate is a publication containing various information. Each article will be published to the main website when completed. On occasion, small typos, grammar, or spelling errors will be caught that were not be found in the initial editing stages. If this occurs, they will be corrected and a new version will be uploaded with the fixes. The date of the document will not change from it's original. It is also possible that larger errors may be found. One of the main purposes of this document is to apply the scientific approach and dissemination of information. For this reason, other additions may be added as feedback from readers is received. \textcolor{Red}{If this is the case, these additions will be added in red lettering and a small note will be added to the title page stating that there was an update with a date attached.} This is done for potential referencing situations. The format of this document is defined below. This copywrite/about page is subject to potentially change in later versions.
	
\vspace{0.5cm}
	
\begin{framed}[.5\textwidth]
	\begin{minipage}{\linewidth}
		\section*{Article Title}
		Each issue contains multiple articles. Each article is it's own thought and may or may not be related to the other articles.
		
		\hspace{1em}\subsection*{Section}
		
		An article may have multiple sections. Each section should be related to the overall article focus but may provide a logical separation between the various elements of the article.
		
		\hspace{1em}\subsubsection*{Sub-section}
		
		Sub-sections are simple dividers that may be used within the various sections. This is to simply provide the reader with shorter parts to read or stopping points throughout. Each article will be concluded with the symbol below.
		
		\sectionEnd
		
		\section*{References}
		The references are provided at the end of each volume for all of the articles it contains. They are in alphabetical order.
	\end{minipage}
\end{framed}


	\vspace{1cm}
		
	The original maintainer of this work is: Antonius Torode. \\		
	The current maintainer of this work is: Antonius Torode. \\
	Chief Editor: Antonius Torode\\		 		
	Published by Antonius Torode. 
	
	\textbf{Hosted at}: https://PeerReviewedPirate.github.io/
	
	
	\vfill
	
	\textbf{Citation:} \\
	\hspace*{2em} Torode, A.\\
	\hspace*{2em} <Article Title>. \\
	\hspace*{2em} Peer Reviewed Pirate. \\
	\hspace*{2em} \DateToUse. \\
	\hspace*{2em} Volume. \Version. \\
	\hspace*{2em} https://PeerReviewedPirate.github.io/
	
	\endgroup
	\clearpage
\twocolumn